---
title: "A comparison of retinal nerve fiber layer thickness in pseudopapilledema and papilledema"
author: "Alexei Marcilio, Merrimack College"
date: "Mar 28th, 2018"
output:
  pdf_document: default
  html_document:
    css: lab.css
    highlight: pygments
    theme: cerulean
  word_document: default
---

```{r echo=FALSE, results='hide', message=FALSE}
require(ggplot2)

require(captioner)

library(ggthemes)
library(dplyr)
library(tidyr)
library(directlabels)
library(ggrepel)
library(extrafont)
library(stringr)
library(Amelia)
library(corrgram)
library(corrplot)
require(gridExtra)
require(ggmosaic)
require(lsr)
library(zoo)
library(caTools)
library(randomForest)
library(e1071)
library(purrr)
library(caret)
library(stats)


figs <- captioner(prefix="Figure")
tbls <- captioner(prefix="Table")

knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = FALSE, dev = "cairo_pdf")

```

\pagenumbering{gobble}

\newpage

\pagenumbering{arabic}

\tableofcontents

\newpage

__Executive Summary__


\newpage

__Introduction__


In clinical practice, a common dilemma is to distinguish pseudopapilledema from pathologic optic disc edema. 
Diagnosis and management of pseudopapilledema and papilledema are vastly different. However, frequently, 
patients with elevated optic discs are almost always referred for neuro-ophthalmologic evaluation and may 
be subjected to neuroimaging, lumbar puncture, and other laboratory studies. It is more challenging to the 
practitioner when patients have papilledema superimposed on pseudopapilledema because patients with 
pseudopapilledema have elevated optic nerve head appearances at baseline. It is important to note that 
papilledema is a serious clinical sign that may represent underlying threatening intracranial pathology. 
Similarly, it is as important to recognize features of pseudopapilledema to avoid unnecessary costly and invasive tests.

Optical coherence tomography (OCT), a non-invasive imaging method, which creates in vivo cross sections of the retina, 
has been a valuable tool and is used to understand some neurological diseases such as optic neuritis, 
multiple sclerosis, spinocerebellar ataxia, neuromyelitis optica, pituitary tumors and Parkinson's disease.
Studies have  compared retinal nerve fiber layer (RNFL) thickness in patients with pseudopapilledema and papilledema.  
The purpose of this study is to evaluate and characterize findings of patients with pseudopapilledema and 
papilledema superimposed pseudopapilledema to help clinicians understand disease progression and resolution 
by describing patients with papilledema superimposed on pseudopapilledema, assessing the upper-limit of RNFL 
thickness in patients with pseudopapilledema, identifying differences in variability of RNFL measurements of 
patients during follow-up, and proposing recommendations on how to approach patients with pseudopapilledema.

Data was collected from 2010-2017, a 6-year period, from the database of patients of the neuro-ophthalmology 
service at New England Eye Center at Tufts Medical Center, Boston Massachusetts
Twenty-four eyes of 12 patients were included in pseudopapilledema group and 14 eyes of 7 patients were 
included in papilledema superimposed on pseudopapilledema group. 
Statistical analysis and machine learning techniques were used to evaluate the the accuracy of predicting 
whether a patient has pseudopapilledema or the more serious papilledema. Such a predictive model could potentially
aid in diagnosis resulting in cost savings and increased patient safety by reducing the need for invasive tests.

##__Results__



__Data Cleaning__


```{r}
# Open the loan_data.csv file and save it as a dataframe called pap.
# Set wd
setwd("/Users/alexeimarcilio/Desktop/portfolio/medical")
pap <- read.csv("pseudopapRNFLdataforR_Tam.csv")
```

Let's check the structure of the data.

```{r}
str(pap)
summary(pap)
```

Features that should be factors are changed. Group, which indicates pseudopapilledema or
papilledema, gender, and eye, which indicates left or right, should all be factors.

```{r}
pap$group <- as.factor(pap$group)
pap$gender <- as.factor(pap$gender)
pap$eye <- as.factor(pap$eye)

```

The column names are changed to make then more meaningful. 

```{r}
colnames(pap) <- c('ID','Group','Gender','AgePresented','DurOfFU','Months',
                   'AvgRNFL','RNFL.I','RNFL.S','RNFL.N','RNFL.T','eye')
```

Gender is indicated as 2 for male and 1 for female. Group 2 indicates Papilledema and 
Group 1 indicates Pseudopapilledema, and eye 1 indicates the left eye, while 2 indicates the 
right eye.  The values for these features will be changed  to make them more meaningful. 
Several small functions are written to accomplish this.

```{r}
# This function will update gender
update_gender <- function(gender){
  gender <- as.character(gender)
  if (gender == "1"){
    return('F')
  }else if (gender == "2" ) {
    return('M')
  }else{
    return(gender)
  }
}
# Now we use sapply to apply the function to all rows
pap$Gender <- sapply(pap$Gender, update_gender)
# change it back to a factor
pap$Gender <- as.factor(pap$Gender)
# Now let's do the same for group
update_group <- function(group){
  group <- as.character(group)
  if (group == "1"){
    return('Pseudo')
  }else if (group == "2" ) {
    return('Pap')
  }else{
    return(group)
  }
}
# Now we use sapply to apply the function to all rows
pap$Group <- sapply(pap$Group, update_group)
# change it back to a factor
pap$Group <- as.factor(pap$Group)
# Now let's do the same for group
update_eye <- function(eye){
  eye <- as.character(eye)
  if (eye == "1"){
    return('Left')
  }else if (eye == "2" ) {
    return('Right')
  }else{
    return(eye)
  }
}
# And finally for eye
pap$eye <- sapply(pap$eye, update_eye)
# change it back to a factor
pap$eye <- as.factor(pap$eye)

```

__Data Exploration__

__Missing Values__

Let's check for missing values. The visits of each patient are recorded by month after the initial
visit which is indicated as month 0. Every subsequent month, from 1 to 66, has a placeholder and this 
accounts for the multiple missing values.

```{r}
 missmap(pap, main='', col = c("black","wheat"),
         legend=FALSE,  x.cex = 0.7, y.labels = NULL, y.at = NULL, rank.order = TRUE) 

```

As these months in which a patient has no visits are not necessary we can remove them. There are no missing
values for the patient demographic data. 

```{r}
pap <- na.omit(pap)

```

After all missing values were removed the missing map chart shows no missing values.

####__Correlations__

The average RNFL is not included here as it's a derived field that will be highly correlated with the other RNFL measures.
Each RNFL measure is moderately correlated to each other in the left eye, but more strongly correlated in the right eye.
RNFL.S is particularly strongly correlated with RNFL.I in the right eye (r$^2$ = 0.8). In the bottom panels long, narrow ellipses represent high correlations while circular ellipses represent low correlations.

```{r}

  pap %>% 
  select(-AvgRNFL, -ID) %>% 
  filter(eye=="Left") %>% 
  corrgram(order=TRUE, lower.panel = panel.ellipse,
         upper.panel = panel.cor, text.panel = panel.txt)

  pap %>% 
  select(-AvgRNFL, -ID) %>% 
  filter(eye=="Left") %>% 
  corrgram(order=TRUE, lower.panel = panel.ellipse,
         upper.panel = panel.cor, text.panel = panel.txt)


```



####__Distributions__

The distributions of the numeric fields were examined. There are a wide range of ages, months, 
and visit durations. The continuous variables of interest show fairly normal unimodal distributions
that are skewed to the right.

```{r}

# pap %>% 
#   select(-ID) %>% 
#   keep(is.numeric) %>% 
#   gather() %>% 
#   ggplot(aes(value)) +
#   facet_wrap(~ key, scales = "free") +
#   geom_histogram(alpha=0.6, color="black", fill="royalblue1") + 
#   theme_wsj(base_family = "serif") +
#   theme(axis.text.x= element_text(size=6, face=1)) +
#   theme(axis.text.y= element_text(size=6, face=1))

pap %>% 
  select(-ID) %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_freqpoly(alpha=1, color="royalblue1", size=1) + xlab("") +
  theme_tufte() +
  theme(text=element_text(size=14, family="Tw Cen MT")) +
  theme(axis.text.x= element_text(size=8, face=1)) +
  theme(axis.text.y= element_text(size=8, face=1)) + 
  xlab("RNFL Thickness")


```

An examination of the distributions of Age, Duration and Month of Visit between Papilledema and
Pseudopapilledema patients was made. The ages of Pseudopapilledema patients are more widely distributed. 
The Papilledema patients are all below 40.The visit duration and months values are similarly distributed
however Papilledema patients are more skewed to the higher values for these measures. 

```{r}

# pap %>% 
#   select(Group, AgePresented, DurOfFU, Months) %>% 
#   gather(key="x", value="value", -Group) %>% 
#   ggplot(aes(value)) + 
#   geom_histogram(aes(fill=Group),binwidth = 5, alpha=0.5, color='black') +
#   facet_wrap(~x, scales = "free") + theme_wsj(base_family = "serif") +
#   theme(axis.text.x= element_text(size=8, face=1)) + xlab("h") +
#   theme(axis.text.y= element_text(size=8, face=1)) + theme(legend.title=element_blank()) + 
#   scale_fill_manual(values=c("orangered1", "royalblue1"), labels=c("Papilledema","Pseudopapilledema")) 

pap %>% 
  select(Group, AgePresented, DurOfFU, Months) %>% 
  gather(key="x", value="value", -Group) %>% 
  ggplot(aes(value)) + 
  geom_freqpoly(aes( color=Group),binwidth = 10, alpha=1, size=1) +
  facet_wrap(~x, scales = "free") + 
  theme_tufte() +
  theme(legend.position="top") +
  theme(axis.text.x= element_text(size=8, face=1)) + xlab("") + 
  theme(text=element_text(size=15, family="Tw Cen MT")) +
  theme(axis.text.y= element_text(size=8, face=1)) + theme(legend.title=element_blank()) + 
  scale_color_manual(values=c("orangered1", "royalblue1"), labels=c("Papilledema","Pseudopapilledema")) 

```


Next we examine the distribution of Average RNFL values between Papilledema and Pseudopapilledema
patients. The average RNFL measurements from Papilledema patients are more highly skewed. Only Papilledema
patients present average RNFL measurement over 200 $\mu$m.

```{r}
#pap %>%
#   ggplot(aes(x=AvgRNFL)) +
#   geom_histogram(aes(fill=Group),binwidth = 10, alpha=0.5, color='black') +
#   theme_wsj(base_family = "serif") +
#   theme(axis.text.x= element_text(size=8, face=1)) +
#   theme(axis.title=element_text(size=10, face=1)) +
#   theme(axis.text.y= element_text(size=8, face=1)) + theme(legend.title=element_blank()) +
#   scale_fill_manual(values=c("orangered1", "royalblue1"), labels=c("Papilledema","Pseudopapilledema"))

pap %>% 
  ggplot(aes(x=AvgRNFL)) +
  geom_freqpoly(aes( color=Group),binwidth = 10, alpha=1, size=1.2) +
  theme(axis.text.x= element_text(size=8, face=1)) +
  theme(axis.title=element_text(size=10, face=1)) 
  theme(text=element_text(size=15, family="Tw Cen MT")) 
  theme(axis.text.y= element_text(size=8, face=1)) + theme(legend.title=element_blank())  
  scale_color_manual(values=c("orangered1", "royalblue1"), labels=c("Papilledema","Pseudopapilledema")) 
  
```

We can examine the distributions of the inferior, superior, nasal and temporal retinal nerve 
fibre layer thicknesses (RNFL.I, RNFL.S, RNFL.N, RNFL.T) in patients with Papilledema vs. those 
with Psuedopapilledema. The distributions of all these measurements show a similar pattern of 
Papilledema values being skewed to higher values. The superior retinal nerve fibre layer thickness
shows one value which appears to be an outlier. 

```{r}

# pap %>% 
#   select(Group, RNFL.I, RNFL.S, RNFL.N, RNFL.T ) %>% 
#   gather(key="x", value="value", -Group) %>% 
#   ggplot(aes(value)) + 
#   geom_histogram(aes(fill=Group),binwidth = 20, alpha=0.6, color='black') +
#   facet_wrap(~x, scales = "free") + 
#   ylab("Count") +
#   xlab("RNFL thickness") +
#   theme_wsj(base_family = "serif") +
#   theme(axis.text.x= element_text(size=8, face=1)) +
#   theme(axis.text.y= element_text(size=8, face=1)) + theme(legend.title=element_blank()) + 
#   scale_fill_manual(values=c("orangered1", "royalblue1"), labels=c("Papilledema","Pseudopapilledema")) 

pap %>% 
  select(Group, RNFL.I, RNFL.S, RNFL.N, RNFL.T ) %>% 
  gather(key="x", value="value", -Group) %>% 
  ggplot(aes(value)) + 
  geom_freqpoly(aes(color=Group),binwidth = 20, alpha=1, size= 1) +
  facet_wrap(~x, scales = "free") + 
  ylab("Count") +
  xlab("RNFL thickness") +
  theme_tufte() +
  theme(legend.position="top") +
  theme(text=element_text(size=10, family="Tw Cen MT")) +
  theme(axis.text.x= element_text(size=8, face=1)) +
  theme(axis.text.y= element_text(size=8, face=1)) + theme(legend.title=element_blank()) + 
  scale_color_manual(values=c("orangered1", "royalblue1"), labels=c("Papilledema","Pseudopapilledema")) 

```

Boxplots better illustrate the difference in variance of the these measures for Papilledema vs. 
Psuedopapilledema patients. The outlier was removed so that the distributions could be better compared
side by side. For each retinal measurement the median value is generally higher for Papilledema patients vs.
Pseudopapilledema. The range of values is also much greater for Papilledema patients.

```{r}

group_names <- c("Left" = "Left Eye", "Right" = "Right Eye", "RNFL.I" = "RNFL.I", 
                 "RNFL.S" = "RNFL.S", "RNFL.N" = "RNFL.N", "RNFL.T" = "RNFL.T")

pap %>% 
  select(Group, eye, RNFL.I, RNFL.S, RNFL.N, RNFL.T ) %>% 
  filter(RNFL.S < 1000) %>% 
  gather(key="x", value="value", -Group, -eye) %>% 
  ggplot(aes(x=Group,y=value, fill=Group)) + 
  geom_boxplot(alpha=.6, color='black', aes(color=Group)) +
  scale_color_manual(values = c("orangered1", "royalblue1")) +
  scale_fill_manual(values=c("orangered1", "royalblue1")) +
  geom_jitter(alpha=.3, aes(color=Group)) +
  facet_grid(eye~x, labeller = as_labeller(group_names)) + 
  theme_tufte() +
  theme(text=element_text(size=14, family="Tw Cen MT")) +
  theme(axis.text.x= element_text(size=9, face=1)) +
  theme(axis.text.y= element_text(size=9, face=1)) + 
  theme(legend.position="none") + ylab("Thickness") + xlab("Papilledema vs. Pseudopapilledema")

```


The trends in average RNFL thickness are examined by charting the values over time for each patient. 
The range in values for average RNFL thickness is much greater for Papilledema patients than 
Pseudopapilledema. The RNFL values of each Pseudopapilledema patient stay within a relatively 
narrow range over time. The one outlier was removed.


```{r}

group_names <- c("Pap" = "Papilledema","Pseudo" = "Pseudopapilledema", "Left" = "Left Eye", "Right" = "Right Eye")

pap  %>% 
  ggplot(aes(x=Months, y=AvgRNFL)) + 
  geom_point(aes(color=factor(ID)), alpha = 1, size =1.2 ) +
  geom_line(aes(color=factor(ID)), alpha=1, size = 1) +
  theme_tufte() +
  theme(text=element_text(size=14, family="Tw Cen MT")) +
  theme(legend.title=element_blank()) +
  theme(legend.position="none") + ylab("Average RNFL") +
  theme(axis.text.x= element_text(size=8, face=1), axis.text = element_text(size=20, face=1)) +
  theme(axis.text.y= element_text(size=8, face=1)) +
  facet_grid(eye ~ Group, labeller = as_labeller(group_names))

```

We can examine the sames trends for each individual RNFL measurement. The pattern is similar to the average RNFL
for each measurement. There's a much greater amount of variation between each measurement over time for the 
Papilledema patients compared to the Pseudopapilledema patients.


```{r fig.align="center"}
group_names <- c("Pap" = "Pap.","Pseudo" = "Pseudo.", 
                 "Left" = "Left Eye", "Right" = "Right Eye",
                 "RNFL.I" = "RNFL.I", "RNFL.S" = "RNFL.S", 
                 "RNFL.N" = "RNFL.N", "RNFL.T" = "RNFL.T")

  pap %>% 
  select(ID, Group, eye, Months, RNFL.I, RNFL.S, RNFL.N, RNFL.T ) %>% 
  filter(RNFL.S < 1000) %>% 
  gather(key="x", value="value", -Group, -eye, -Months, -ID) %>% 
  ggplot(aes(x=Months, y=value)) + 
  geom_point(aes(color=factor(ID)), alpha = 1, size =.8 ) +
  geom_line(aes(color=factor(ID)), alpha=.7, size = .9) +
  theme_tufte() + 
  theme(text=element_text(size=13, family="Tw Cen MT")) +
  theme(legend.title=element_blank()) +
  theme(legend.position="none") + ylab("Average RNFL") +
  theme(axis.text.x= element_text(size=9, face=1), axis.text = element_text(size=10, face=1)) +
  theme(axis.text.y= element_text(size=9, face=1)) +
  facet_grid(eye ~ Group ~ x, labeller = as_labeller(group_names))
```

##__Feature Engineering__

* Roll up measures to the patient level - for example left_RNFL.I will be a feature - one feature for each 
eye retinal length measure for all months (average?), also max min perhaps, number of visits, total time between first
and last visit etc. Might require some programming.

* t-tests on these features.

* random forest - even though there are 19 Pseudo and 7 Pap maybe this would be useful?

* Logistic regression?

* A spreadsheet or application in which a doctor can enter these measures and a probability of Pap is returned?

#__Parametric Tests__

```{r }

x.Psuedo <- 
  pap %>% 
  select(ID, AvgRNFL, Group, eye, Months) %>% 
  filter(eye == 'Left', Months == 0, Group == 'Pseudo') %>% 
  select(AvgRNFL)

y.Pap <- 
  pap %>% 
  select(ID, AvgRNFL, Group, eye, Months) %>% 
  filter(eye == 'Left', Months == 0, Group == 'Pap') %>% 
  select(AvgRNFL)



t.test(x.Psuedo, y.Pap)


```
```{r}
pap %>% 
  select(ID, Group) %>% 
  filter(Group == 'Pseudo') %>% 
n_distinct()
```


#__Machine Learning__

\newpage

#__References__

